% --- Abbreviations --- %
\newglossary*{abbr}{Abbreviations}

\newacronym[type=abbr]{xa}{xA}{expected Assists}
\newacronym[type=abbr]{xg}{xG}{expected Goals}
\newacronym[type=abbr]{xt}{xT}{expected Threat}

% --- Parameters --- %
\newglossary*{param}{Parameters}

\newglossaryentry{param:bdp}{type=param,
    name={\textit{BDP}},
    description={Buildup Disruption Percentage}}

\newglossaryentry{param:daxt}{type=param,
    name={\textit{DxAT}},
    description={Defensive Action expected Threat}}

\newglossaryentry{param:psr}{type=param,
    name={\textit{PSR}},
    description={Pressing Success Rate}}

\newglossaryentry{param:vpep}{type=param,
    name={\textit{VPEP}},
    description={Valuing Pressure Events by Estimating Probabilities}}

\newglossaryentry{param:xg}{type=param,
    name={\textit{xG}},
    description={expected Goals}}

\newglossaryentry{param:xt}{type=param,
    name={\textit{xT}},
    description={expected Threat}}

% --- Nomenclature --- %
\newglossary*{nomen}{Nomenclature}

\newglossaryentry{nomen:pwin}{type=nomen,
    name={$P_{\text{win}}$},
    description={Probability of winning the ball}}

\newglossaryentry{nomen:pgoal}{type=nomen,
    name={$P_{\text{goal/chance}}$},
    description={Probability of conceding a goal-scoring chance}}

\newglossaryentry{nomen:t}{type=nomen,
    name={$t$},
    description={Time variable used in pressing analysis}}

\newglossaryentry{nomen:x}{type=nomen,
    name={$x$},
    description={Spatial location on the field (\(x\)-coordinate}}
    
\newglossaryentry{nomen:y}{type=nomen,
    name={$y$},
    description={Spatial location on the field (\(y\)-coordinate}}

\newglossaryentry{nomen:w}{type=nomen,
    name={$w$},
    description={Weight or significance of pressing action}}

\newglossaryentry{nomen:mu}{type=nomen,
    name={$\mu$},
    description={Mean (average pressing intensity or event probability)}}

\newglossaryentry{nomen:sigma}{type=nomen,
    name={$\sigma$},
    description={Variance (to represent variability in pressing effectiveness)}}

% --- Terminology --- %
\newglossary*{term}{Terminology}

\newglossaryentry{term:ballrec}{type=term,
    name={ball recovery},
    description={the act of regaining possession of the ball, typically following a press or interception.}}

\newglossaryentry{term:batch}{type=term,
    name={batch},
    description={subset of training data used in one iteration.}}

\newglossaryentry{term:compactness}{type=term,
    name={compactness},
    description={the degree to which a team stays close together, minimising space for opponents to exploit.}}

\newglossaryentry{term:counter}{type=term,
    name={counter-attacking},
    description={a playing philosophy focused on quickly attacking after regaining possession.}}

\newglossaryentry{term:counterpress}{type=term,
    name={counter-press},
    description={the act of pressing immediately after losing possession, often within a few seconds of having lost the ball.}}

\newglossaryentry{term:defstrat}{type=term,
    name={defensive strategies},
    description={these include overall approaches like deep-defending (maintaining a compact shape near the goal) and high-press (applying pressure high up the pitch).}}

\newglossaryentry{term:deftrans}{type=term,
    name={defensive transition},
    description={the phase in which a team shifts from attack to defence, often involving immediate pressing to regain possession.}}

\newglossaryentry{term:dropout}{type=term,
    name={dropout},
    description={optimise by randomly dropping out neurons during training.}}

\newglossaryentry{term:epoch}{type=term,
    name={epoch},
    description={one pass through the entire training dataset.}}

\newglossaryentry{term:xa}{type=term,
    name={expected Assists (xA)},
    description={a measure of the likelihood that a pass will result in a goal.}}

\newglossaryentry{term:xg}{type=term,
    name={expected Goals (xG)},
    description={a metric that quantifies the likelihood of a shot resulting in a goal, based on historical shot data.}}

\newglossaryentry{term:xt}{type=term,
    name={expected Threat (xT)},
    description={a metric designed to quantify the danger posed by having possession of the ball in a specific location on the pitch (or resulting from a specific action).}}

\newglossaryentry{term:gegenpress}{type=term,
    name={gegenpressing},
    description={high-intensity pressing tactic that emphasises immediate pressure after losing possession}}